Hyperplane Elimination for Quickly Enumerating Local Optima

### 2015-09-15 Hyperplane Enumeration/main.tex
Fixed largest Adjacent NKq size

\begin{abstract}
A common tool for understanding combinatorial problem landscapes
is to examine the properties of its local optima. Unfortunately,
exhaustive methods for finding local optima are exceptionally time
consuming and are therefore limited to very small problem sizes.
We propose a method for exploiting generalized problem structure
to significantly speed up this process by skipping hyperplanes
which cannot contain local optima. In most of our experiments
this hyperplane elimination allows enumeration time to scale with
the total number of local optima instead of the total landscape size.
We prove optimality for linear and $k$-bounded problems, and provide
empirical evidence of optimality on NKq Landscapes and Ising Spin Glasses.
Furthermore, we are able to extend this method to finding only solutions
which cannot be improving by flipping $r$ or fewer bits. In some cases
increasing $r$ actually reduces total runtime by drastically reducing the
number of local optima. While previous methods were limited to
problems with at most 18 or 30 bits, hyperplane elimination can enumerate the same problems with
75 bits. When using higher $r$ values this bound can be pushed to 200 bit problems
on average enumerated in under 6 minutes.

@@

Here we introduce a method for finding all local optima for the same problems using
up to 75 bits. The cornerstone of this method is the identification
of hyperplanes that cannot contain any local optima, allowing large
portions of the search space to be skipped during enumeration.
The techniques developed here can be applied to the generalized problem class of Mk Landscapes,
which contains many real world and benchmark combinatorial problems.

@@

This reduction in growth complexity translates into a significant increase in the size of problems
which can be enumerated. For instance, using the previous best enumeration methods~\cite{ochoa:2015:crossovernetworks}
the largest Adjacent NKq problem size which can be enumerated in 4 hours is 34 bits. Given the same time
Hyper-Reorder can enumerate problems with 75 bits. Using the regression model, we estimate solving
a 75 bit problem using Gray-Box would require 230 million years. Similarly, Hyper-Reorder's largest
Random NKq size of 69 bits would require Gray-Box 3 million years. These larger sizes make for
much better approximations of landscapes interesting to search methods.

>>>>

\begin{abstract}
A common tool for understanding combinatorial problem landscapes
is to examine the properties of its local optima. Unfortunately,
exhaustive methods for finding local optima are exceptionally time
consuming and are therefore limited to very small problem sizes.
We propose a method for exploiting generalized problem structure
to significantly speed up this process by skipping hyperplanes
which cannot contain local optima. In most of our experiments
this hyperplane elimination allows enumeration time to scale with
the total number of local optima instead of the total landscape size.
We prove optimality for linear and $k$-bounded problems, and provide
empirical evidence of optimality on NKq Landscapes and Ising Spin Glasses.
Furthermore, we are able to extend this method to finding only solutions
which cannot be improving by flipping $r$ or fewer bits. In some cases
increasing $r$ actually reduces total runtime by drastically reducing the
number of local optima. While previous methods were limited to
problems with at most 18 or 30 bits, hyperplane elimination can enumerate the same problems with
77 bits. When using higher $r$ values this bound can be pushed to 200 bit problems
on average enumerated in under 6 minutes.

@@

Here we introduce a method for finding all local optima for the same problems using
up to 77 bits. The cornerstone of this method is the identification
of hyperplanes that cannot contain any local optima, allowing large
portions of the search space to be skipped during enumeration.
The techniques developed here can be applied to the generalized problem class of Mk Landscapes,
which contains many real world and benchmark combinatorial problems.

@@

This reduction in growth complexity translates into a significant increase in the size of problems
which can be enumerated. For instance, using the previous best enumeration methods~\cite{ochoa:2015:crossovernetworks}
the largest Adjacent NKq problem size which can be enumerated in 4 hours is 34 bits. Given the same time
Hyper-Reorder can enumerate problems with 77 bits. Using the regression model, we estimate solving
a 77 bit problem using Gray-Box would require 900 million years. Similarly, Hyper-Reorder's largest
Random NKq size of 69 bits would require Gray-Box 3 million years. These larger sizes make for
much better approximations of landscapes interesting to search methods.
